---
title: "Handling non-nested hierarchies with `rtauargus`"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    social: menu
    source_code: embed
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "Z:/SDC/OutilsConfidentialite/protect_tables_examples/")
library(flexdashboard)
library(dplyr)
library(rtauargus)
options(
  rtauargus.tauargus_exe =
    "Y:/Logiciels/TauArgus/TauArgus4.2.3/TauArgus.exe"
    # "C:/Users/tmm7an/Documents/SDC/TauArgus/TauArgus4.2.3/TauArgus.exe"
)
# library(data.table)
# library(ggraph)
# library(igraph)
library(flextable)
library(collapsibleTree)
load("data/turnover_act_size.rda")
load("data/turnover_act_cj.rda")
load("data/activity_corr_table.rda")
table_1 <- turnover_act_size
table_2 <- turnover_act_cj
hier_main <- activity_corr_table %>% 
  filter(!A21 %in% c("U","V","W","X","Y","Z")) %>% 
  filter(substr(A88,2,2) != "X")
hier_alt <- hier_main %>%
  filter(A21 %in% LETTERS[4:13]) %>%
  mutate(A10 = "D_TO_M") %>%
  unique()
```



Overview
=====================================

Row {data-height=200}
-----------------------------------------------------------------------

### **The release**

We'd like to release two tables:  

1. The turnover of companies broken down by their `ACTIVITY` (business sector) and their `SIZE`.  
2. The turnover of companies broken down by their `ACTIVITY` (business sector) and their `CJ` (legal category).  

The `ACTIVITY` variable is broken down into three levels of the NACE (or equivalent) hierarchy.
But, we'd like also to publish an additional aggregation: th sum of the D to M levels of the NACE. 
We call it `D_TO_M`.

### **Confidentiality rules**

- *frequency rule* threshold = 3  
- *dominance rule* (1,85%)  
- *Interval protection level* = 10%  
- *singleton rule*


Row {data-height=800}<!-- Row {.tabset .tabset-fade data-height=800} -->
-----------------------------------------------------------------------

### **TABLE 1**

```{r}
DT::datatable(
  table_1,
  rownames = FALSE,  
  options = list(
    dom = 'tp',scrollY=FALSE,pageLength=15
  )
) %>% 
  DT::formatRound(3:5, digits=0, dec.mark = ".", mark = " ")
```


### **TABLE 2**

```{r}
DT::datatable(
  table_2,
  rownames = FALSE, 
  options = list(dom = 'tp',scrollY=FALSE,pageLength=15)
) %>% 
  DT::formatRound(3:5, digits=0, dec.mark = ".", mark = " ")
```


Hierarchies
=====================================

Row
-----------------------------------------------------------------------

### **Activity: the main hierarchy**

```{r}
hier_main_long <- tibble(
  root = "Total", leaf = hier_main$A10 %>% unique(), level = 1
) %>% 
  bind_rows(
    hier_main %>%
      select(root = A10, leaf = A21) %>% 
      unique() %>% 
      mutate(level = 2)
  ) %>% 
  bind_rows(
    hier_main %>% 
      select(root = A21, leaf = A88) %>% 
      unique() %>% 
      mutate(level = 3)
  )
```


```{r}
collapsibleTree(hier_main, c("A10", "A21", "A88"), root = "Total")
```


### Activity: an alternative hierarchy

```{r}
collapsibleTree(hier_alt, c("A21", "A88"), root = "D_to_M", fill = "steelblue")
```


Building the list of tables
=====================================

Row {data-height=350}
-----------------------------------------------------------------------

### **Analysis**

1. The two original tables share some common cells on the margins of, respectively, `SIZE` and `CJ` levels. So, they can't be protected independently.  

2. Adding the `D_TO_M` total introduces a non-nested hierarchy issue

$\Rightarrow$ We need to split each original table into two tables:

a. `table_1` = `ACTIVITY` $\times$ `SIZE` with the levels of the main `ACTIVITY` hierarchy
b. `table_1_alt` = `ACTIVITY` $\times$ `SIZE` with the levels of the alternative `D_TO_M` hierarchy
c. `table_2` = `ACTIVITY` $\times$ `SIZE` with the levels of the main `ACTIVITY` hierarchy
d. `table_2_alt` = `ACTIVITY` $\times$ `SIZE` with the levels of the alternative `D_TO_M` hierarchy

$\Rightarrow$ The protection has to be done on these four tables.

Row  {data-height=650}
-----------------------------------------------------------------------

### **TABLES 1**

```{r}
table_1_alt <- table_1 %>%
  filter(
    ACTIVITY %in% LETTERS[4:13]
  ) %>%
  bind_rows(
    table_1 %>%
      filter(
        ACTIVITY %in% LETTERS[4:13]
      ) %>%
      group_by(SIZE) %>%
      summarise(N_OBS = sum(N_OBS), TOT = sum(TOT), MAX = max(MAX)) %>%
      mutate(ACTIVITY = "D_TO_M")
  )
```

#### **`ACTIVITY` levels in `table_1`**

```{r}
table_1 %>% pull(ACTIVITY) %>% unique()
```

#### **`ACTIVITY` levels in `table_1_alt`**

```{r}
table_1_alt %>% pull(ACTIVITY) %>% unique()
```


### **TABLES 2**

```{r}
table_2_alt <- table_2 %>%
  filter(
    ACTIVITY %in% LETTERS[4:13]
  ) %>%
  bind_rows(
    table_2 %>%
      filter(
        ACTIVITY %in% LETTERS[4:13]
      ) %>%
      group_by(CJ) %>%
      summarise(N_OBS = sum(N_OBS), TOT = sum(TOT), MAX = max(MAX)) %>%
      mutate(ACTIVITY = "D_TO_M")
  )
# str(table_2_alt)
```


#### **`ACTIVITY` levels in `table_2`**

```{r}
table_2 %>% pull(ACTIVITY) %>% unique()
```

#### **`ACTIVITY` levels in `table_2_alt`**

```{r}
table_2_alt %>% pull(ACTIVITY) %>% unique()
```


Preparation of the data
=====================================

Row
-----------------------------------------------------------------------

### **Preparation of hrc files**

The `rtauargus::write_hrc2()` function converts a correspondence table into a 
proper `.hrc` file.

```{r echo=TRUE}
hrc_main <-  hier_main %>%
  write_hrc2(file_name = "hrc/hrc_main")

hrc_alt <- hier_alt %>%
  select(A21) %>%
  unique() %>% 
  write_hrc2(file_name = "hrc/hrc_alt")
```

The structure of the input is the following one:

```{r}
hier_main %>% str()
```


The produced `.hrc` file is then:

```{r}
read.table(hrc_main) %>% select(` `=V1)
```


### **Preparation of the list of tables to protect**

All the tables have to be included in a named list. We also have to detect cells 
that don't respect the primary rules (frequency and dominance rules). For this purpose, 
we create a boolean variable (`is_secret_prim`) that takes `TRUE` if a cell doesn't respect a primary secret rule and then has to be suppress.

```{r echo=TRUE}

list_tables <- 
  # Create the list with our four tables
  list(
    t1 = table_1,
    t1_alt = table_1_alt,
    t2 = table_2,
    t2_alt = table_2_alt
  ) %>%
  # apply detection of primary cells in every table
  purrr::map(
    \(df){
      df %>%
        mutate(
          #frequency rule
          is_secret_freq = N_OBS > 0 & N_OBS < 3,
          #dominance rule
          is_secret_dom = ifelse(MAX == 0, FALSE, MAX/TOT>0.85),
          #primary status
          is_secret_prim = is_secret_freq | is_secret_dom
        )
    }
  )
```

The structure of the tables is then:

```{r}
str(list_tables)
```



Implementation with `rtauargus`
=====================================

Row 
-----------------------------------------------------------------------

### **Code** 


```{r echo=TRUE, eval=FALSE}
res <- tab_multi_manager(
    # The list of tables to protect 
    list_tables = list_tables,
    # The explanatory vars of each table of the list
    list_explanatory_vars = 
      purrr::map(list_tables, \(df) df %>%  select(1:2) %>% names()),
    # The hrc file of each hierarchical variable
    hrc = c(ACTIVITY = hrc_main),
    # In case of alternative hierarchies in a given table: mention them in alt_hrc
    alt_hrc = list(
      t1_alt = c(ACTIVITY = hrc_alt), t2_alt = c(ACTIVITY = hrc_alt)
    ),
    # The directory where to save the tau-argus files
    dir_name = "outputs/",
    # response variable
    value = "TOT",
    # frequency variable
    freq = "N_OBS",
    # var indicating the primary secret satus
    secret_var = "is_secret_prim",
    # total code for each variable
    totcode =  "Total",
    # In case of alternative totals in a given table, mention them in alt-totcode
    alt_totcode = list(
      t1_alt = c(ACTIVITY = "D_TO_M"), t2_alt = c(ACTIVITY = "D_TO_M")
    )
  )
```



### **Outputs**

```{r}
shiny::actionButton(
  "launch",
  label = "Launch protection"
)
```

```{r}
protect_res <- eventReactive(input$launch, {
  
  res <- tab_multi_manager(
    # The list of tables to protect 
    list_tables = list_tables,
    # The explanatory vars of each table of the list
    list_explanatory_vars = 
      purrr::map(list_tables, \(df) df %>%  select(1:2) %>% names()),
    # The hrc file of each hierarchical variable
    hrc = c(ACTIVITY = hrc_main),
    # In case of alternative hierarchies in a given table: mention them in alt_hrc
    alt_hrc = list(
      t1_alt = c(ACTIVITY = hrc_alt), t2_alt = c(ACTIVITY = hrc_alt)
    ),
    # The directory where to save the tau-argus files
    dir_name = "outputs/",
    # response variable
    value = "TOT",
    # frequency variable
    freq = "N_OBS",
    # var indicating the primary secret satus
    secret_var = "is_secret_prim",
    # total code for each variable
    totcode =  "Total",
    # In case of alternative totals in a given table, mention them in alt-totcode
    alt_totcode = list(
      t1_alt = c(ACTIVITY = "D_TO_M"), t2_alt = c(ACTIVITY = "D_TO_M")
    )
  )
  
})

summary_res <- reactive({
  
  protect_res() %>% 
    purrr::imap_dfr(
      function(df, name){
        df %>% 
          rename_with(~"is_secret_final", last_col()) %>% 
          mutate(final_status = case_when(
          is_secret_freq ~ "freq",
          is_secret_dom ~ "dom",
          is_secret_final ~ "sec.",
          TRUE ~ "Valid"
        )) %>% 
          group_by(final_status) %>% 
          summarise(
            n_cells = n(),
            val_cells = sum(TOT),
            .groups = "drop"
          ) %>% 
          mutate(pc_nb_cells = n_cells/sum(n_cells)*100,
                 pc_val_cells = val_cells/sum(val_cells)*100) %>% 
          mutate(table = name) %>% 
          select(table, final_status, n_cells, pc_nb_cells, val_cells, pc_val_cells)
      }
    )
  
})
```

```{r}
shiny::renderPrint({
  a <- protect_res()
})
```


```{r}
loc_journal <- file.path(getwd(), "outputs/journal.txt")
```


All the files have been recorded within the directory mentionned in the `dir_name` argument.
A journal, indicating all the steps of the protection process, is also written. This journal isn't
very easy to use in the case of big tables.


### **Summary**


```{r}
renderUI({
  options(digits=1)
  summary_res() %>%
    select(-val_cells) %>% 
    flextable() %>%
    colformat_int(
      big.mark = " ",
      na_str = "na") %>% 
    colformat_num(
      j=4:5,big.mark = " ",dec.mark=".", digits = 2, na_str = "na"
    ) %>% 
    theme_booktabs() %>%
    autofit() %>%
    htmltools_value()
})
```


Control of coherence
=====================================

### Confusion Matrix 1


```{r}
conf_matrices <- reactive({
  
  mat1 <- protect_res()$t1 %>% 
    select(1,2,is_secret_1 = last_col()) %>% 
    inner_join(
      protect_res()$t1_alt %>% 
    select(1,2,is_secret_1_alt = last_col())
    ) %>% 
    group_by(is_secret_1, is_secret_1_alt) %>% 
    count() %>% 
    tidyr::pivot_wider(names_from=is_secret_1_alt, values_from=n, values_fill = 0)
  
  mat2 <- protect_res()$t2 %>% 
    select(1,2,is_secret_2 = last_col()) %>% 
    inner_join(
      protect_res()$t2_alt %>% 
    select(1,2,is_secret_2_alt = last_col())
    )%>% 
    group_by(is_secret_2, is_secret_2_alt) %>% 
    count() %>% 
    tidyr::pivot_wider(names_from=is_secret_2_alt, values_from=n, values_fill = 0)
  
  mat3 <- protect_res()$t1 %>% 
    filter(SIZE == "Total") %>% 
    select(1,is_secret_1 = last_col()) %>% 
    inner_join(
      protect_res()$t2 %>% 
        filter(CJ == "Total") %>% 
    select(1,is_secret_2 = last_col())
    ) %>% 
    group_by(is_secret_1, is_secret_2) %>% 
    count() %>% 
    tidyr::pivot_wider(names_from=is_secret_2, values_from=n, values_fill = 0)
  
  return(list(mat1,mat2, mat3))
})
```


The confusion matrix below compares the final status (suppressed or not) of the common cells of
`table_1` and `table_1_alt`.

```{r}
renderUI({
  options(digits=1)
  conf_matrices()[[1]] %>%
    # rename(~"Secret status in table_1", is_secret_1) %>% 
    flextable() %>%
    add_header_row(
      top = TRUE, 
      values = c("", "Secret status in table_1_alt"), 
      colwidths = c(1, 2)
    ) %>% 
    bg(i = 1, j = 3, bg = "orange") %>% 
    bg(i = 2, j = 2, bg = "orange") %>% 
    theme_booktabs() %>%
    autofit() %>%
    htmltools_value()
})
```


### Confusion matrix 2

The confusion matrix below compares the final status (suppressed or not) of the common cells of
`table_2` and `table_2_alt`.

```{r}
renderUI({
  options(digits=1)
  conf_matrices()[[2]] %>%
    # rename(~"Secret status in table_2", is_secret_1) %>% 
    flextable() %>%
    add_header_row(
      top = TRUE, 
      values = c("", "Secret status in table_2_alt"), 
      colwidths = c(1, 2)
    ) %>% 
    bg(i = 1, j = 3, bg = "orange") %>% 
    bg(i = 2, j = 2, bg = "orange") %>% 
    theme_booktabs() %>%
    autofit() %>%
    htmltools_value()
})
```


### Confusion matrix 3

The confusion matrix below compares the final status (suppressed or not) of the common cells of
`table_1` and `table_2`.

```{r}
renderUI({
  options(digits=1)
  conf_matrices()[[3]] %>%
    # rename(~"Secret status in table_1", is_secret_1) %>% 
    flextable() %>%
    add_header_row(
      top = TRUE, 
      values = c("", "Secret status in table_2"), 
      colwidths = c(1, 2)
    ) %>% 
    bg(i = 1, j = 3, bg = "orange") %>% 
    bg(i = 2, j = 2, bg = "orange") %>% 
    theme_booktabs() %>%
    autofit() %>%
    htmltools_value()
})
```

